---
title: "report"
author: "Jose Munoz Angulo, Vivien Marcault, Baptiste Rigondaud"
date: "February 26, 2018"
output: html_document
---

## Question 1

The map is produced with this code:

//TODO: insert here

The map is created by extracting all unique indigenous population along with its latitude and longitude. By the means of the maps library we draw the map with the populations in its location, using 9 different colors and three different symbols.

## Question 2

Performing the linear regression:

// TODO: insert code here

R can not invert the matrix to solve te linear regression problem, because of the size of the matrix. Thus, it explains why the lieanr regression summary is full of Not A Number. In order to have a correct model, we could take a subset of the predictors using this command:

//TODO: insert code here

In the following questions, we are going to use PCA.

## Question 3

### a)
This method consist into diagonalizing the covariance matrix in order to find its maximum eigenvalues. These values correspond to the components of maximum variance. Therefore if we make a projection of the data on the factorspace generated by the corresponding eigenvectors, we preserve the maximum information that we can have from the original data in this dimension. 

### b)
TODO: We use the scale parameter in order to normalize the variance
```{r}
pcaNAm2 = prcomp(NAm2[,-c(1:8)], scale=TRUE)
```

### c)
```{r}
caxes=c(1,2)
plot(pcaNAm2$x[,caxes],col="white")
for (i in 1:npop) {
  print(names[i])
  lines(pcaNAm2$x[which(NAm2[,3]==names[i]),caxes],
        type="p",col=colPalette[i],pch=pch[i])
}
legend("bottomleft",legend=names,col=colPalette,lty=-
         1,pch=pch,cex=.75,ncol=3,lwd=2)
```

The graph maps the weight of each genetic samples regarding the two first components of the PCA. We can see that these two first components gathers all the members of a population, but gathers also all the population around the same area.
The 5th and th 6th axes separate more the populations (Pima, Karitiana) compared to the two first axes.

### d)
In order to have the percentages, we use the command:
//TODO: insert code here
The first two principal components capture $3.4%$ of the variance.

## Question 4

### a)
// TODO: insert code here

Comparing with the first map, we can see that the population are well located and well centered around the expected position based on training data. But the map illustrate too optimistically the ability to locate people according to their genetic material. In fact we tested our model with the training data.

### b)
